---
title: "Untitled"
author: "Eric Skuse"
date: "4/11/2020"
output: html_document
---

# Adult Income Data Analysis

**Data Understanding & Preparation** 
 
The data used for our business problem was retrieved from Kaggle: [adult income](https://www.kaggle.com/serpilturanyksel/adult-income)
 
This dataset includes 48842 observations (rows) and with 15 variables (columns). Each observation in the data set represents an individual adult. Within each observation, variables include:Age  Work class  Education  Marital status  Occupation  Relationship  Race  Gender  Capital gain  Capital loss  Hours per week  Native country  Salary  

```{r}
rm(list = ls())

library(readr)
library(tidyverse)
library(ggplot2)
library(readr)
setwd("~/Documents/GitHub/DSBI")
adult11 <- read_csv("Data/adult11.csv", col_types = cols(X16 = col_skip(), X17 = col_skip()), na = "?")
View(adult11)
```

## Data clean up
```{r}
#Remove observations with missing data and assign variables as factors
adult <- na.omit(adult11)
adult$workclass <- factor(adult$workclass)
adult$education <- factor(adult$education)
adult$`marital-status` <- factor(adult$`marital-status`)
adult$occupation <- factor(adult$occupation)
adult$relationship <- factor(adult$relationship)
adult$race <- factor(adult$race)
adult$gender <- factor(adult$gender)
adult$salary <- factor(adult$salary)
```

```{r}
#Check column names
colnames(adult)

#change column names from dash to underscore
colnames(adult)<- c("age","workclass","fnlwgt","education","education_num","marital_status","occupation","relationship","race","gender","capital_gain","capital_loss","hours_per_week","native_country","salary")

summary(adult)
```
```{r}
#To make sure we have a reasonable number of observations to model with, check how many people in the data set made capital gains more than 0
adult%>%
  filter(capital_loss>0)

#How many made losses more than 0?
adult%>%
  filter(capital_gain>0)%>%
  arrange(desc(`capital_gain`))
```

```{r}
#Check for distribution in capital gain and loss data: it looks like lots of people made 99,999 in capital gains (it must be the cap). But no one made 40-100k in gains? Seems improbable. Maybe the entries for 99,999 in capital gains are mistakes.
adult%>%
  ggplot(aes(x=age,y=capital_gain,color=salary)) +
  geom_point(shape=1)

adult%>%
  ggplot(aes(x=age,y=capital_loss,color=salary)) +
  geom_point(shape=1)
```
```{r}
#another look at the distribution of the capital gain and capital loss data points
adult%>%
  filter(capital_gain>0)%>%
    ggplot(aes(x=capital_gain))+
    geom_histogram(binwidth = 200)

adult%>%
  filter(capital_loss>0)%>%
    ggplot(aes(x=capital_loss))+
    geom_histogram(binwidth = 50)
```

**Remove outliers**
```{r}
#take out all the observations with 99999 capital gains, since they may skew the results
adult[adult == 99999] <- NA
adult <- na.omit(adult)

#check to see if obbservations were removed
adult%>%
  filter(capital_gain>0)%>%
    ggplot(aes(x=capital_gain))+
    geom_histogram(binwidth = 200)
```


```{r}
#people with more salary have higher capital gains
adult%>%
  filter(capital_gain>0)%>%
  ggplot(aes(x=salary,y=capital_gain))+
  geom_boxplot()

adult%>%
  filter(capital_loss>0)%>%
  ggplot(aes(x=salary,y=capital_loss))+
  geom_boxplot()
```

**Establish a target variable**
```{r}
#Need to create some new columns to model our target variable, i.e. who would like to play in the stock market and be receptive to marketing efforts from our financial services firm? Let's start with a binary Yes / no variable for people who either gained or lost more than $1,000. We'll mark them as play=yes

adult<-adult%>%
  mutate(play=as.factor(ifelse(capital_loss >= 500 | capital_gain >= 500,"Yes","No")))

adult%>%
  filter(play=="Yes")
```

# Modeling

## Logistic regression model
```{r}
#Education, salary, marital status, and gender are predictors with significance
logit_model<-glm(play~gender+salary+education+race+workclass+marital_status+occupation+relationship,
                 family="binomial", 
                 data=adult)                   
summary(logit_model)
```
```{r}
#Re-run the model with only gender, education and salary as predictors
logit_model<-glm(play~gender+salary+education+marital_status,
                 family="binomial", 
                 data=adult)                   
summary(logit_model)
```

## Use of the model to predict
```{r}
adult$log_odd<-predict(logit_model)                         # get predicted log odds (default)
adult$logit_pred_prob<-predict(logit_model,type="response") # get predicted probabilities

adult%>%
  select("play","capital_gain","capital_loss","log_odd","logit_pred_prob")
```

```{r}
#With the predicted probabilities, you can sort people by the predicted probability of having significant capital gains or losses
adult%>%
  arrange(desc(logit_pred_prob))%>%
  select("play", "capital_gain", "capital_loss", "salary", "logit_pred_prob")
```

```{r}
#Use a different cut-off for class prediction using `ifelse()`.
adult$logit_pred_class<-ifelse(adult$logit_pred_prob>0.3,"Yes","No")

adult%>%
  arrange(desc(logit_pred_prob))%>%
  select("play", "capital_gain", "capital_loss", "salary", "logit_pred_prob", "logit_pred_class")
```

## Model validation
```{r}
set.seed(3)   # set a random seed 
index <- sample(nrow(adult), nrow(adult)*0.2) # random selection of indices. 
test <- adult[index,]       # save 20% as a test dataset
training <- adult[-index,]   # save the rest as a training set
```

```{r}
#run the model on training data only
logit_model<-glm(play~gender+salary+education+marital_status,
                 family="binomial", 
                 data=training)                   
summary(logit_model)
```

```{r}
#apply model to test data
logit_model_test<-glm(play~gender+salary+education+marital_status,
                 family="binomial", 
                 data=test)                   
summary(logit_model_test)

```

```{r}
test$lm_log_odd<-predict(logit_model_test,test)
test$lm_pred_prob<-predict(logit_model_test,test,type="response")

test$lm_bp_pred_class=ifelse(test$lm_pred_prob>0.3,"Yes","No")

table(test$lm_bp_pred_class,test$play, dnn=c("predicted","actual")) 

table(test$lm_bp_pred_class==test$play)
7580/8998         #accuarcy of the model = 84%

head(test)
```

```{r}
#ROC curve
library(pROC)
lm_roc<-roc(test$play,test$lm_pred_prob,auc=TRUE)
plot(lm_roc,print.auc=TRUE,col="blue")
```

