---
title: "Adult Income Data Analysis"
author: "Eric Skuse"
date: "4/11/2020"
output: html_document
---

# Data understanding & preparation
 
The data used for our business problem was retrieved from Kaggle: [adult income](https://www.kaggle.com/serpilturanyksel/adult-income)
 
This dataset includes 48842 observations (rows) and with 15 variables (columns). Each observation in the data set represents an individual adult. Within each observation, variables include:Age  Work class  Education  Marital status  Occupation  Relationship  Race  Gender  Capital gain  Capital loss  Hours per week  Native country  Salary  

```{r}
rm(list = ls())
library(readr)
library(tidyverse)
library(ggplot2)
library(readr)
setwd("~/Documents/GitHub/DSBI")
adult11 <- read_csv("Data/adult11.csv", col_types = cols(X16 = col_skip(), X17 = col_skip()), na = "?")
View(adult11)
```

## Data clean up

```{r}
#Remove observations with missing data and assign variables as factors
adult <- na.omit(adult11)
adult$workclass <- factor(adult$workclass)
adult$education <- factor(adult$education)
adult$`marital-status` <- factor(adult$`marital-status`)
adult$occupation <- factor(adult$occupation)
adult$relationship <- factor(adult$relationship)
adult$race <- factor(adult$race)
adult$gender <- factor(adult$gender)
adult$salary <- factor(adult$salary)
```

```{r}
#Check column names
colnames(adult)

#change column names from dash to underscore
colnames(adult)<- c("age","workclass","fnlwgt","education","education_num","marital_status","occupation","relationship","race","gender","capital_gain","capital_loss","hours_per_week","native_country","salary")

summary(adult)
```
```{r}
#To make sure we have a reasonable number of observations to model with, check how many people in the data set made capital gains more than 0
adult%>%
  filter(capital_loss>0)

#How many made losses more than 0?
adult%>%
  filter(capital_gain>0)%>%
  arrange(desc(`capital_gain`))
```

```{r}
#Check for distribution in capital gain and loss data: it looks like lots of people made 99,999 in capital gains (it must be the cap). But no one made 40-100k in gains? Seems improbable. Maybe the entries for 99,999 in capital gains are mistakes.
adult%>%
  ggplot(aes(x=age,y=capital_gain,color=salary)) +
  geom_point(shape=1)

adult%>%
  ggplot(aes(x=age,y=capital_loss,color=salary)) +
  geom_point(shape=1)

adult%>%
  ggplot(aes(x=age,y=hours_per_week,color=salary)) +
  geom_point(shape=1)
```
```{r}
#another look at the distribution of the capital gain and capital loss data points
adult%>%
  filter(capital_gain>0)%>%
    ggplot(aes(x=capital_gain))+
    geom_histogram(binwidth = 200)

adult%>%
  filter(capital_loss>0)%>%
    ggplot(aes(x=capital_loss))+
    geom_histogram(binwidth = 50)

adult%>%
  filter(hours_per_week>0)%>%
    ggplot(aes(x=hours_per_week))+
    geom_histogram(binwidth = 1)
```

## Remove outliers

```{r}
#take out all the observations with 99999 capital gains, since they may skew the results
adult[adult == 99999] <- NA
adult <- na.omit(adult)

#check to see if obbservations were removed
adult%>%
  filter(capital_gain>0)%>%
    ggplot(aes(x=capital_gain))+
    geom_histogram(binwidth = 200)
```


```{r}
#people with more salary have higher capital gains
adult%>%
  filter(capital_gain>0)%>%
  ggplot(aes(x=salary,y=capital_gain))+
  geom_boxplot()

adult%>%
  filter(capital_loss>0)%>%
  ggplot(aes(x=salary,y=capital_loss))+
  geom_boxplot()

adult%>%
  filter(hours_per_week>0)%>%
  ggplot(aes(x=salary,y=hours_per_week))+
  geom_boxplot()
```

# Modeling

## Logistic regression model
```{r}
logit_model<-glm(salary~gender+age+capital_gain+capital_loss+education+race+workclass+marital_status+occupation+relationship+native_country+hours_per_week,
                 family="binomial", 
                 data=adult)                   
summary(logit_model)
```
```{r}
#Re-run the model with only gender, education, and race as predictors
logit_model<-glm(salary~gender+race+education,
                 family="binomial", 
                 data=adult)                   
summary(logit_model)
```

# Use of the model to predict
```{r}
adult$log_odd<-predict.glm(logit_model)                         # get predicted log odds (default)
adult$logit_pred_prob<-predict.glm(logit_model,type="response") # get predicted probabilities

adult%>%
  select("salary","race","gender","education","log_odd","logit_pred_prob")
```

```{r}
#With the predicted probabilities, you can sort people by the predicted probability of having salaries higher than $50K per year
adult%>%
  arrange(desc(logit_pred_prob))%>%
  select("salary", "race","gender", "education", "gender", "logit_pred_prob")
```

```{r}
#Use a different cut-off for class prediction using `ifelse()`.
adult$logit_pred_class<-ifelse(adult$logit_pred_prob>0.5,"Yes","No")

adult%>%
  arrange(desc(logit_pred_prob))%>%
  select("salary", "race", "gender", "education", "logit_pred_prob", "logit_pred_class")
```

# Model validation
```{r}
set.seed(2)   # set a random seed 
index <- sample(nrow(adult), nrow(adult)*0.2) # random selection of indices. 
test <- adult[index,]       # save 20% as a test dataset
training <- adult[-index,]   # save the rest as a training set
```

```{r}
#run the model on training data only
logit_model<-glm(salary~gender+race+education,
                 family="binomial", 
                 data=training)                  
summary(logit_model)
```

```{r}
#apply model to test data
logit_model_test<-glm(salary~gender+race+education,
                 family="binomial", 
                 data=test)                     
summary(logit_model_test)

```

```{r}
test$lm_log_odd<-predict.glm(logit_model_test,test)
test$lm_pred_prob<-predict.glm(logit_model_test,test,type="response")

test$lm_bp_pred_class=ifelse(test$lm_pred_prob>0.5,">50K","<=50K")

table(test$lm_bp_pred_class,test$salary, dnn=c("predicted","actual")) 

table(test$lm_bp_pred_class==test$salary)

head(test)
```

## ROC curve

```{r}
library(pROC)
lm_roc<-roc(test$salary,test$lm_pred_prob,auc=TRUE)
plot(lm_roc,print.auc=TRUE,col="blue")
```